Lab 2 Final Report: COVID-19 Case Rate vs Population Demographics and Mask Policy

### 2020-12-04 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Updated regression table and discussion

The research question aims to measure the COVID-19 case rate per 100,000 residents within each state based on the make-up of the population, as well as the policy decisions that were or were not put into place in order to combat the rise of COVID-19 cases. In this sense, the aim is to examine how dependent the COVID-19 case rate is on that of which cannot be controlled (i.e., population demographics), as well as that of which can be controlled to a certain degree (i.e., implementation of policies to attempt to combat case rate). With this information, one could suggest whether or not the proliferation of COVID-19 seems to be related to either or both controllable and uncontrollable variables, and to the level at which the effect of these variables differ help. The modeling goal of this research question will be one of description, and will be broken up into three phases of investigation

@@

Firstly, it is important to note that given each state is treated as a unique data point, the sample contains 51 data points. Although this size meets the general rule that an adequate sample size is 30 data points or more, as the analysis begins to factor in the wide range of potential population distributions and demographics within the states, it is clear that there is large variation within the sample. For example, the population density can have large variability depending on the population concentration among a few large areas as well as the level of uninhabited or sparsely inhabited land within a state, while the demographics of population can depend on a wide range of things such as regional factors and employment opportunities.

@@

```{r}  
# Percentage variables are re-scaled to raters per 100k

@@

```{r, warning=FALSE, message=FALSE}
ggplot(data = df,
  mapping = aes(x= case_rate_100k)) +
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,6000,600)) +
  labs(title =  "Histogram of COVID-19 Case Rate per 100,000", 
       x = "COVID-19 case rate per 100,000", y = 'Count') +
  scale_x_continuous(breaks=seq(0, 6000, 600))
```

@@

ggplot(data = df,
  mapping = aes(x= senior_per_100k)) +
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(10000,21000,1000)) +
  labs(title = 'Histogram of Seniors per 100,000', 
  x = 'Population of seniors per 100,000', y = 'Count') + 
  scale_x_continuous(breaks=seq(10000,22000,1000))
```

@@

```{r message=FALSE}
df %>% 
  ggplot(aes(senior_per_100k, case_rate_100k)) + 
  geom_smooth(se = FALSE) +
  geom_point() +   
  labs(
    title = 'COVID-19 Case Rate due to Senior Population', 
    x = 'Population of senior per 100,000', 
    y = 'COVID case rate per 100,000'
  )
```

@@

### Sam - @Frank, this was my previous analysis of the POPULATION DENSITY variable, feel free to use this

@@

Although there were several versions of models created for model 2 and 3, through the analysis of the appropriate variables respective to each model a single model version was selected to be included in the regression table. This was determined based on factors of significance when investigating the effect relationship of these variables on the outcome variable, which is detailed in section 2.

For the regression table below, model 1, model 2-b, and model 3-a were selected as the most significant models to include.

@@

stargazer(model1, model2, model3, type = "text",
          se = list(se.model3),
          star.cutoffs = c(0.05, 0.01, 0.001), 
          title = "Table 1: The relationship between COVID case rate and population density")

```

@@

# 5-1: Urban vs Rural Population Ratio

@@

# 5-2: Median Income

@@

# 5-3: Gender Ratio

@@

# 5-4: Mask Use Among Population

@@

# 5-5: GDP Percentage of Tourism

>>>>

The research question aims to measure the COVID-19 case rate per 100,000 residents within each state based on the make-up of the population, as well as the policy decisions that were or were not put into place in order to combat the rise of COVID-19 cases. In this sense, the aim is to examine how dependent the COVID-19 case rate is on that of which cannot be controlled (i.e., population demographics), as well as that of which can be controlled to a certain degree (i.e., implementation of policies to attempt to combat case rate) by each state. With this information, one could suggest whether or not the proliferation of COVID-19 within a state seems to be related to either or both controllable and uncontrollable variables, and to the level at which the effect of these variables differ help. The modeling goal of this research question will be one of description, and will be broken up into three phases of investigation

@@

Firstly, it is important to note that given each state is treated as a unique data point, the sample contains 51 data points. Although this size meets the general rule that an adequate sample size is 30 data points or more, as the analysis begins to factor in the wide range of potential population distributions and demographics within the states, it is clear that there is large variation within the sample. For example, the population density can have large variability depending on the population concentration among a few large areas as well as the level of uninhabited or sparsely inhabited land within a state, while the demographics of population can depend on a wide range of things such as regional factors and employment opportunities. This point is further discussed in the IID assumption addressed in section 3 of this report.

@@

```{r}  
# Percentage variables are re-scaled to rates per 100k

@@

```{r, warning=FALSE, message=FALSE}
ggplot(data = df,
  mapping = aes(x= case_rate_100k)) +
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,6000,600)) +
  labs(title =  "Histogram of COVID-19 Case Rate", 
       x = "COVID-19 case rate per 100,000", y = 'Count') +
  scale_x_continuous(breaks=seq(0, 6000, 600))
```

@@

ggplot(data = df,
  mapping = aes(x= senior_per_100k)) +
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(10000,21000,1000)) +
  labs(title = 'Histogram of Seniors Distribution', 
  x = 'Number of seniors per 100,000 people', y = 'Count') + 
  scale_x_continuous(breaks=seq(10000,22000,1000))
```

@@

```{r message=FALSE}
df %>% 
  ggplot(aes(senior_per_100k, case_rate_100k)) + 
  geom_smooth(se = FALSE) +
  geom_point() +   
  labs(
    title = 'COVID-19 Case Rate due to Senior Population', 
    x = 'Number of seniors per 100,000', 
    y = 'COVID case rate per 100,000'
  )
```

@@

### @Frank, this was my previous analysis of the POPULATION DENSITY variable, feel free to use any of this

@@

With the analysis for the selected variables, the three corresponding models, and the respective diagnostics of the CLM assumptions complete, the models can then be included for comparison in a regression table. Model 1, model 2, and model 3 are included in the regression table below,.

@@

stargazer(model1, model2, model3, type = "text",
          se = list(se.model3),
          star.cutoffs = c(0.05, 0.01, 0.001), 
          title = "Table 1: The effect of population demographics and mask policy on COVID-19 case rate")
```

@@

### 5-1: Urban vs Rural Population Ratio

@@

### 5-2: Median Income

@@

### 5-3: Gender Ratio

@@

### 5-4: Mask Use Among Population

@@

### 5-5: GDP Percentage of Tourism
